Ripple Consensus Review

### 2015-04-26 paper.tex
wrote up some attacks

\documentclass{article}
\usepackage{url}
\usepackage{mathtools}
\usepackage{tikz}
\usepackage{multirow}
\usetikzlibrary{arrows,positioning}
\begin{document}
\title{Ripple Consensus Review}
\author{Peter Todd}
\date{FIXME}
\maketitle

@@

\begin{table}
    \begin{tabular}{c|c|c|c|}
        ~ & \multicolumn{3}{c}{Majority UNL \%} \\
        \multirow{3}{*}{You} & $100\%$ & $50\%$ & $10\%$ \\ \hline
        $100\%$ & Consensus & DoS & DoS \\ \hline
        $50\%$ & DoS & DoS & D.S. \\ \hline
        $10\%$ & D.S & D.S & D.S. \\
    \end{tabular}
\end{table}

Claims from ripple: consensus protocol\cite{ripple-consensus-paper} will maintain correctness so long as $f
\le (n-1)/5$ where $f$ is the number of Byzantian failures. However default UNL
list is just $5$ entries, thus $f \le (5-1)/5 = 0.8$ - no failures can be
tolerated.

Put antoher way, you only need to compromise $20\%$ of the nodes in the UNL to
censor a transactions/freeze funds. "failure to agree is agreement to defer"

Is there incentive to use anything but the default UNL? Remember that if you
choose a UNL more than $80\%$ different than others, you can be attacked on
that basis; your full node determines validity of transactions anyway. e.g. the
"add 1000 validators to UNL list" suggestion.

mixing structure: if tx A is rejected, dependent txs are all rejected; will
likely soon result in situation where you can't use different UNL

The Ripple consensus algorithm starts with a Unique Node List.

what happens with UNL majority is not reached?

test: add extra entries to UNL list, see if ledgers close, if they don't think we know threshold

\section{Attacks}

reference c3 criteria?

@@

1) exploit consensus failure.

DoS: Cost \$0, Scope: global, Duration: hours
Theft: Cost \$1k, Scope: varies, Duration: hours

@@

\subsection{Scalability}

DoS: Cost \$100k, Scope: global, Duration: weeks

simply produce a lot of txs, paid for via fees, either covertely or overtly.
System will fail, guaranteed.

@@

>>>>

\documentclass{article}
\usepackage{url}
\usepackage{mathtools}
\usepackage{tikz}
% \usepackage{multirow}
\usetikzlibrary{arrows,positioning}
\begin{document}
\title{Ripple Consensus Review}
\author{Peter Todd}
\date{FIXME}
\maketitle

@@

\begin{table}
    \begin{tabular}{c|c|c|c|}
        ~ & \multicolumn{3}{c}{Majority UNL \%} \\
        % \multirow{3}{*}{You} & $100\%$ & $50\%$ & $10\%$ \\ \hline
        $100\%$ & Consensus & DoS & DoS \\ \hline
        $50\%$ & DoS & DoS & Doublespend \\ \hline
        $10\%$ & Doublespend & Doublespend & Doublespend \\
    \end{tabular}
\end{table}

\section{Attack scenarios}

To evaluate potential attacks on the Ripple network we look at three factors:

\begin{itemize}

    \item \textbf{Type:} What is the effect of the attack? Denial of Service
          and/or theft? Or is the attack mostly useful in preparation for another
          type of attack?

    \item \textbf{Cost:} What is a probable minimum cost to carry out the attack, for
          the people able to do the attack? For instance, a technical attack
          that can be launched by ``bored teenagers'' will have an essentially
          zero cost while a long-term rewrite of a proof-of-work blockchain has
          a reasonably well-defined cost.

    \item \textbf{Scope:} Who is affected by the attack? Is this a targetted attack,
          mainly affecting a one or two targets? An attack with broad impact? Or
          an attack with global impact on all Ripple users?

    \item \textbf{Duration:} How long will it take before the attack is
          neutralized?  This may be a matter of hours to write and distribute a
          simple software patch\footnote{The author personally stopped the
          CVE-2013-4627 attack on the Bitcoin network in a few hours of rushed
          work in the middle of the night.}, to indefinitely long for some
          attacks where recovery is impossible.

    \item \textbf{Probability:} How likely is it for the Ripple network to be
          attacked this way?

\end{itemize}

We do \emph{not} try to determine the monetary impact of the attack, because we
are unable to predict how the Ripple network will be used. For instance, while
a double-spend attack may have the same basic impact on a $\$100/\text{day}$
small business as it would on a $\$1,000,000,000/\text{day}$ multinational, the
latter could lose orders of magnitude more money for what it at a technical
level basically the same attack. A closely issue with attempting to predict
monetary losses is that at some point losses can be sufficiently high that they
create situations where social consensus can decide that the losses simply must
be reversed. For instance, Vericoin community chose to do a hard-fork of their
alt-currency to undo the impact of a major theft.\cite{coindesk-vericoin}

@@

The attacker exploits a difference in behavior between different
implementations/versions of the Ripple protocol. The result of the attack can
be a simple denial of service due to the Ripple network being unable to process
transactions, or with more sophistication, the attacker can fork the Ripple
ledger/blockchain and use the fork to doublespend. The attack is stopped by
identifying it, and then distributing software patches. Provided the Ripple
network is well monitored consensus split attacks can be stopped fairly
quickly; the March 2013 Bitcoin chain fork\cite{bip50} was resolved in a matter
of hours by contacting mining pools. A similar response could happen on Ripple
by contacting validating nodes commonly used in UNLs.\footnote{Another reason
to use the UNL provided Ripple Labs.}

Note that a consensus split may also happen by accident; prior consensus splits
in cryptocurrency systems have almost always been accidental. Equally, that
consensus splits can happen by accident clearly shows that the minimum cost to
perform this attack is zero. However, due to the limited number of consensus
bugs available to exploit in a given (set of) versions of the Ripple protocol
in use and the fast response time possible to consensus splits the potential of
this class of attack for long-term attacks on the network is limited.

Risk factors:

\begin{itemize}

    \item The main implementation of the Ripple protocol, \emph{rippled}, does
        not cleanly separate the consensus-critical part of the codebase from
        the non-consensus-critical part.\footnote{The consensus-critical
            portions of Bitcoin Core codebase are currently being separated
            into a stand-alone \emph{libconsensus} library.}

    \item The Ripple protocol itself is extremely complex, with many different
          types of transactions, and all functionality being implemented directly
          in the protocol itself.

    \item Improvements to Ripple frequently require changing the
          consensus-critical codebase rather than user software. For instance
          while on Bitcoin the implementation of multisig was possible without
          modification to the protocol\cite{bip19} in Ripple the lack of
          extension capabilities such as scripting require a consensus-critical
          change.\cite{ripple-wiki-multisign}\footnote{P2SH was \emph{not}
          required for multisig to be implemented.}

\end{itemize}

DoS: Cost \$0, Scope: global, Duration: hours, Prob: high
Theft: Cost \$1k, Scope: varies, Duration: hours, Prob: high

@@

\subsection{Transaction Flood}

DoS: Cost depends, Scope: broad, Duration: depends

The attacker creates large numbers of transactions - either covertely or
overtly - until the Ripple network fails due to overload, or more likely,
Ripple users are priced out of their ability to use the Ripple network.
Transactions pay a minimum fee per transaction by destroying XRP; the fee per
transaction is set by a vote between all validators on the UNL list. In the
short to medium term such an attack would quickly drive up that fee. A
succesful attack requires deep pockets as the attacker must outspend other
Ripple users.

Risk factors:

\begin{itemize}

    \item Ripple Labs plans\cite{cointelegraph-ripple-aml} to require verified
        user identification to use the Ripple network in the near future.  When
        implemented this will make it easy to determine who is doing a covert
        transaction flood attack and restrict their access to the network.

    \item The global consensus ledger has inherently poor $O(n^2)$ scaling;
        even in the absense of a malicious attack the poor scalability of the
        ledger is a threat to the viability of Ripple.

    \item A clever attacker wishing to overtly attack the Ripple network can
        masquerade their traffic as legit economic activity; from the point of
        view of Ripple Labs - a major XRP owner - such an attack may not even
        be considered an attack!\footnote{Similar to the debates in the Bitcoin
        ecosystem about whether or not fee-paying uses of the Bitcoin
        blockchain for non-Bitcoin-denominated transactions constitute
        ``attacks''.}

    \item Attacking the poor scalability of Ripple is highly synergistic with
        attacks depending on the high centralization of the Ripple network as
        even ``unsuccesful'' scalability attacks drive further centralization.
        For instance, if the response to a scalability attack is for Ripple
        Labs to beef up the validation nodes they control, and users stop
        running full nodes themselves, the network is more vulnerable to any
        attacks on those centrally managed validation nodes.

\end{itemize}

@@

\subsection{Validator Fear, Uncertainty, and Doubt}

make people afraid to run validating nodes

illegal data
